### 2021-07-06 rad-icfp-2021/complexity.tex
expanding complexity section

We follow a simple model similar to the one in \cite{griewank2008evaluating}.
We assume the cost is divided into 4 elementary measures, being the number of MOVES, ADDS, MULTS, and NLOPS.
MOVES assumes a flat memory and represents moving a fixed size information  (e.g. 64 bits). 
ADDS represents the number of additions, 
MULTS the number of multiplications, 
and NLOPS the number of elementary non linear operations like !cos!, !exp!.

@@

\begin{itemize}
    \item $\cost(*)=(3,0,1,0)$
    \item $\cost(+)=(3,1,0,0)$
    \item $\cost(c)=(1,0,0,0)$
    \item $\cost($!sin!$)=(2,0,0,1)$
\end{itemize}

TODO: extend cost array
TODO: extend cost compositional way

@@

\begin{proposition}
    TODO: cost $\directD{\rho}{\Gamma}{Y}$. 
\end{proposition}

As we can see, the only problem is that too many variables are being moved, 
due to the continuation carrying and copying $n$ variables at each step.
This is exactly the overhead that our inlining and forward substitution removes.

\begin{proposition}
    TODO: complexity gain using our optim.
\end{proposition}

>>>>

We follow a simple model similar to the one in \cite{griewank2008evaluating}.
We assume the cost is divided into $4$ elementary measures, being the number of MOVES, ADDS, MULTS, and NLOPS.
MOVES assumes a flat memory and represents moving a fixed size information  (e.g. 64 bits). 
ADDS represents the number of additions, 
MULTS the number of multiplications, 
and NLOPS the number of elementary non-linear operations like !cos! or !exp!.

@@

\begin{tabular}{ll}
    $\cost(*)=(3,0,1,0)$ & $\cost(+)=(3,1,0,0)$\\
    $\cost(c)=(1,0,0,0)$ & $\cost($!sin!$)=(2,0,0,1)$
\end{tabular}

For simplicity, we will not address any parallelism in our cost model. 
So the cost of !map (x.e)! on an array of size $n$ will be $n*\cost$(!e!).
The cost function extends compositionally to the whole language in the obvious way.
For instance we have !$\cost$(let x=e$_1$ in e$_2$)!= !$\cost$(e$_1$)+$\cost$(e$_2$)!.
TODO: for let binding and same in map, do I need to add some MOVES?

@@

\begin{proposition}
    $\cost(\directD{\rho}{\Gamma}{Y}($!e!$))\leq$ $K\cost$(!e!) + O((0,0,0,n*|!e!|)) 
\end{proposition}

\begin{proof}
    TODO: routinely by induction and inspection. be specific with K? 
    One problem still is for map and reduce because they call twice the gradient of !e$_1$!.
\end{proof}

As we can see, the only problem is that too many variables are being moved, and this is
due to the continuation carrying and copying $n$ variables at each step.
This overhead is precisely what that our inlining and forward substitution removes, 
as was examplified in Section~\ref{sub:Partial evaluation and optimization}.
